%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%
\abstract{
Concrete is a highly complex heterogeneous material.
Not only are the effective mechanical properties highly dependent on various factors, as for example aggregate type, water content or additives, but the properties change over time and the evolution of properties is in turn dependent on the specific mix design of the concrete.
This is already a challenge in it self when only dealing with concrete properties.
To ... something about structural design... connecting them...
By applying .... stochastic methods, the quality of the data can be estimated.
Automated workflow to simplify addition of additional data points and general reproducibility.

By considering the broader implications of concrete mixture optimization and GWP reduction, this research contributes to the advancement of sustainable construction practices.

In summary, this paper introduces a holistic optimization procedure that integrates concrete mixture optimization with the structural design of precast concrete elements, aiming to achieve structurally efficient elements with minimized environmental impact. 
\ET{Abstract needs to be written. 
	This is just a placeholder with some thoughts, probably best to wait till results etc are done.}}

\keywords{performance oriented design, stochastic optimization, precast concrete, mix design}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle

